---
title: "HW1"
author: "Christina Chen"
date: "2026-09-20"
output: github_document
---

```{r setup, echo = F, message = F}
library(tidyverse)
```

## Problem 1
Load the dataset:
```{r load}
data("penguins", package = "palmerpenguins")
```

### Data Description
The `penguins` dataset has information on the penguins' species, island, bill length (mm), bill depth (mm), flipper length(mm), body mass (g), sex, and year. 
The dataset has `r nrow(penguins)` observations and `r ncol(penguins)` variables. 
The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm.

### Scatterplot

```{r plot, warning = F}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguins_scatterplot.png")
```


## Problem 2

Creating data frame:
```{r}
hw1_df = tibble(
  samp = rnorm(10),
  samppos = samp > 0,
  vec_char = c("laundry", "cooking", "dusting", "scrubbing", "mopping", "sweeping", "dishes", "tidying", "ironing", "mending"),
  vec_factor = factor(c("medium", "hard", "easy", "hard", "medium", "medium", "easy", "medium", "easy", "hard"))
)


```





